Composable Generalized Consensus

### 2016-01-09 paper/intro.tex
Small changes in the intro

Surprisingly, at a first glance, many SMR algorithms seem to share a common structure which would indicate that previous formal analysis could be reused. 
That it because, in many cases, researchers obtained a new SMR algorithm by optimizing an existing SMR algorithm for a particular use case.
For example, Fast Paxos~\cite{Lamport06FastPaxos} adds fast rounds to MultiPaxos~\cite{lamport2001paxos} to reduce the latency under low contention, 
Disk Paxos~\cite{GafniLamport03DiskPaxos} optimizes Paxos for systems where disk nodes and processor nodes are distinct, 
Chain Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability} organizes the replicas of Paxos in a ring to distribute commands at a higher throughput but at the expense of latency, 
FGGC~\cite{SutraShapiro11FastGenuineGeneralizedConsensus} optimizes Generalized Paxos\cite{Lamport05GeneralizeConsensus} to reduce latency when the system contains a reliable and fast leader node, 
Ring Paxos~\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol} optimizes the latency and throughput of Paxos when network multicast is available, 
Multicoordinated Paxos reduces the performance impact of a leader fault in MultiPaxos by implementing a distributed leader, 
Cheap Paxos~\cite{LamportMassa04CheapPaxos} optimizes Paxos to decrease the load on some replicas in case failures are sufficiently separated in time,
Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine} optimizes MultiPaxos for wide-area networks by rotating the role of leader among all replicas, etc.
Close scrutiny reveals that most SMR algorithms can be seen as variations and optimizations around a few core ideas such as the round structure introduced by Paxos~\cite{Lamport98ParttimeParliament}, quorum systems~\cite{GuerraouiVukolic10RefinedQuorumSystems,Lamport06FastPaxos}, and agreement up to commutativity using command-structure sets~\cite{Lamport05GeneralizeConsensus}.

@@

To simplify the task of devising new SMR algorithms we propose a framework for
building reusable SMR modules.  
For example, in \cref{}, we augment MultiPaxos with our composition interface and prove that it satisfies Composable Generalized Consensus (CGC), our correctness condition for modules.
We then build an SMR module whose structure is similar to the one of the fast rounds of Fast Paxos and prove that it satisfies CGC\@.
Our Fast Paxos module has lower latency than our Paxos module when there is low contention among clients, but its performance drops below that of Paxos when contention increases. 
Finally, we assemble unmodified our MultiPaxos SMR module and our Fast Paxos SMR module in an SMR algorithm that dynamically switch from one to the other depending on the contention level observed in the system, therefore getting the best of both. 
Crucially, we obtain the correctness of our composite SMR algorithm for free: our generic composition theorem and the correctness of our two SMR modules taken individually imply that the combination of the MultiPaxos SMR module and the Fast Paxos SMR module is a correct SMR algorithm.
Later, in \cref{}, we present three other SMR modules that can be combined the MultiPaxos SMR module and the Fast Paxos SMR module in a composite SMR algorithm that can switch between 5 different kinds of modules, according to the system's operating conditions.

A module in our framework is an SMR algorithm
augmented with \emph{composition interface} and satisfying our correctness condition, called Composable Generalized Consensus (CGC).
The composition interface consists of an \emph{abort interface} for aborting a module's execution and passing
the baton to another module and with a matching \emph{init interface} for taking
over the execution of an aborting module. Two modules are composed by passing
the outputs produced by the abort interface of the first module, called abort
values, as inputs to the init interface of the second module, called init values
in the context of the second module.
A module must satisfy a correctness condition that we call Composable
Generalized Consensus (abbreviated CGC). Any correct module is a safe
Generalized Consensus~\cite{Lamport05GeneralizeConsensus} algorithm (i.e.\ a kind of SMR algorithm), but the progress guarantees may vary between
modules: CGC only requires a module to make progress executing commands or to
abort its execution. CGC also relates the init values, learned values (learned values represent the local state-machine state of individual replicas), and abort values in order for our composition theorem to hold. 
%Note that Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and can offer important performance advantages in practice.

@@

Compared to related work, described in \cref{sec:related}, our key contributions are the following.
\begin{itemize}
    \item We provide tool support for designing CGC modules in the form of two TLA+~\cite{Lamport02SpecifyingSystems} specifications simplifying the design and model-checking of SMR algorithms based on Quorum Systems~\cite{GuerraouiVukolic10RefinedQuorumSystems} and demonstrate their use by designing CGC modules inspired by Paxos, Fast Paxos, and Primary-Backup replication.
    \item Our framework is based on Generalized Consensus, a generalization of the consensus problem 
        that allows optimizing the execution of commuting commands, a common practical case.
\end{itemize}
These two contributions set this paper apart from related work, notably the Abstract framework~\cite{GuerraouiETAL10Next700BftProtocols} and the Speculative Linearizability framework~\cite{GuerraouiKuncakLosa12SpeculativeLinearizability}, both described in \cref{sec:related}.
Our TLA+ specifications can be found at \url{http://losa.fr/cgc}.
Moreover, we have also formalized the CGC correctness condition and proved the composition theorem in Isabelle/HOL\@, and the Isabelle theories are available at the same URL\@.

>>>>

Surprisingly, at a first glance, many SMR algorithms seem to share a
common structure which would indicate that previous formal analysis could
be reused. That is because, in many cases, researchers obtained a new SMR
algorithm by optimizing an existing SMR algorithm for a particular use
case. For example, Fast Paxos~\cite{Lamport06FastPaxos} adds fast rounds
to MultiPaxos~\cite{lamport2001paxos} to reduce the latency under low
contention, Disk Paxos~\cite{GafniLamport03DiskPaxos} optimizes Paxos
for systems where disk nodes and processor nodes are distinct, Chain
Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability} 
organizes the replicas of Paxos in a ring to distribute
commands at a higher throughput but at the expense of latency,
FGGC~\cite{SutraShapiro11FastGenuineGeneralizedConsensus} optimizes
Generalized Paxos\cite{Lamport05GeneralizeConsensus} to reduce latency
when the system contains a reliable and fast leader node, Ring
Paxos~\cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol}
optimizes the latency and throughput of Paxos when network multicast is
available, Multicoordinated Paxos reduces the performance impact of a
leader fault in MultiPaxos by implementing a distributed leader, Cheap
Paxos~\cite{LamportMassa04CheapPaxos} optimizes Paxos to decrease the load
on some replicas in case failures are sufficiently separated in time,
Mencius~\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine} 
optimizes MultiPaxos for wide-area networks by rotating the role of leader
among all replicas, etc. Close scrutiny reveals that most SMR algorithms can
be seen as variations and optimizations around a few core ideas such as the
round structure introduced by Paxos~\cite{Lamport98ParttimeParliament}, quorum
systems~\cite{GuerraouiVukolic10RefinedQuorumSystems,Lamport06FastPaxos},
and agreement up to commutativity using command-structure
sets~\cite{Lamport05GeneralizeConsensus}.

@@

To simplify the task of devising new SMR algorithms we propose a framework
for building reusable SMR modules. For example, in \cref{}, we augment
MultiPaxos with our composition interface and prove that it satisfies Composable
Generalized Consensus (CGC), our correctness condition for modules. We then
build an SMR module whose structure is similar to the one of the fast rounds of
Fast Paxos and prove that it satisfies CGC\@. Our Fast Paxos module has lower
latency than our Paxos module when there is low contention among clients, but
its performance drops below that of Paxos when contention increases. Finally,
we assemble unmodified our MultiPaxos SMR module and our Fast Paxos SMR module
in an SMR algorithm that dynamically switch from one to the other depending on
the contention level observed in the system, therefore getting the best of both.
Crucially, we obtain the correctness of our composite SMR algorithm for free:
our generic composition theorem and the correctness of our two SMR modules taken
individually imply that the combination of the MultiPaxos SMR module and the
Fast Paxos SMR module is a correct SMR algorithm. Later, in \cref{}, we present
three other SMR modules that can be combined the MultiPaxos SMR module and the
Fast Paxos SMR module in a composite SMR algorithm that can switch between 5
different kinds of SMR modules, according to the system's operating conditions.

An SMR module is an SMR algorithm augmented with \emph{composition interface}
and satisfying our correctness condition, called Composable Generalized
Consensus (CGC). The composition interface consists of an \emph{abort interface}
for aborting an SMR module's execution and passing the baton to another
SMR module and with a matching \emph{init interface} for taking over the
execution of an aborting SMR module. Two SMR modules are composed by passing
the outputs produced by the abort interface of the first SMR module, called
abort values, as inputs to the init interface of the second SMR module,
called init values in the context of the second SMR module. A SMR module
must satisfy a correctness condition that we call Composable Generalized
Consensus (abbreviated CGC). Any correct SMR module is a safe Generalized
Consensus~\cite{Lamport05GeneralizeConsensus} algorithm (i.e.\ a kind of SMR
algorithm), but the progress guarantees may vary between SMR modules: CGC only
requires a SMR module to make progress executing commands or to abort its
execution. CGC also relates the init values, learned values (learned values
represent the local state-machine state of individual replicas), and abort
values in order for our composition theorem to hold.
%Note that Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and can offer important performance advantages in practice.

@@

Compared to related work, described in \cref{sec:related}, our key contributions are the following.
\begin{itemize}
    \item We present the first framework to allowing to build Generalized Consensus algorithms our of independent modules.
    \item We provide tool support for designing CGC modules in the form of two TLA+~\cite{Lamport02SpecifyingSystems} specifications simplifying the design and model-checking of SMR algorithms based on Quorum Systems~\cite{GuerraouiVukolic10RefinedQuorumSystems} and demonstrate their use by designing CGC modules inspired by Paxos, Fast Paxos, and Primary-Backup replication.
\end{itemize}
These two contributions set this paper apart from related work, notably the Abstract framework~\cite{GuerraouiETAL10Next700BftProtocols} and the Speculative Linearizability framework~\cite{GuerraouiKuncakLosa12SpeculativeLinearizability}, both described in \cref{sec:related}.
Our TLA+ specifications can be found at \url{http://losa.fr/cgc}.
Moreover, we have also formalized the CGC correctness condition and proved the composition theorem in Isabelle/HOL\@, and the Isabelle theories are available at the same URL\@.
